Understanding Parameterized Quantum Circuit Learning for Quantum Chemical Applications

### 2025-04-03 manuscript/main.tex
Thu  3 Apr 2025 16:59:44 EDT

It should also be noted that the majority of the $t_{2}$-amplitudes will be centered near zero since the $t^{ab}_{ij}$ tensor will contain many small contributions.
Models that exclude small amplitudes, such as those found in Ref\cite{pathirage_exploration_2024}, could be used for increased performance but are out of the scope of this work.

@@

Depth is not always better!
Molecular representations specifically for QML
Distributed QC to incorporate more features
Noiseless simulation is costly and does not offer the desired accuracy for BSE49 or DDCC

DDCC could be a useful dataset to benchmark PQC models since it is trivial to perform classically, yet hard for PQCs...

>>>>

\textcolor{blue}{DDCC Discussion points}
\begin{itemize}
	\item Using multiple feature encoding qubits\cite{suzuki_predicting_2020} is unfeasible with DDCC due to the number of samples
	\item DDCC is a \textit{trivial} task for classical ML (long time to train QML)
	\item It should also be noted that the majority of the $t_{2}$-amplitudes will be centered near zero since the $t^{ab}_{ij}$ tensor will contain many small contributions. Models that exclude small amplitudes, such as those found in Ref\cite{pathirage_exploration_2024}, could be used for increased performance but are out of the scope of this work.
\end{itemize}

@@

PQCs for the BSE49 and DDCC datasets do not exhibit an advantage concerning model accuracy, the number of training points, or time to train the models
Poor performance on the BSE49 dataset could be explained by the feature space reduction, we are compressing the feature space from 2048 to 5 features (structure-based ML reps. require more features than physics-based reps.)
While the physics-based feature set of DDCC outperforms the structure-based feature set of BSE49, DDCC requires too many device executions to be useful on the current generation of devices
For practitioners of quantum chemistry, who work on machine learning-related tasks, like Viki and me, several options exist:
Use hybrid classical-quantum machine learning approaches to exploit entanglement on the QPU, while using a CPU/GPU/TPU for efficient classical tasks
Classical machine learning assisted quantum computing (like I have proposed for a DDCC-LUCJ project for two undergraduates for summer 2025)
Using distributed QML to maintain short circuit depths, while increasing the number of features
Explore molecular representations designed for QML (e.g., BSE49 is a classic-in/quantum-out case)
Stick to quantum-in/quantum-out approaches (e.g., DDCC falls into this category, but could this be reformulated to better exploit the QPU?)
